Learning and comparing functional connectomes across subjects

### 2013-01-01 paper.tex
prediction modeling more

Predictive models are validated using one or more iterations of cross
validation, which involves training and testing phases.  A cross validation
cycle begins by seperating the available data into two independent subsets:
training data and testing data.  Training involves learning a predictive model
from the training data. During testing the model is applied to the testing data
to predict corresponding values of the target variable.  Prediction accuracy is
then assessed by comparing these predictions to the observed values of the
target variable. In order to make

A general application of predictive modeling can be broken down into feature
extraction, model selection, feature selection, and model validation, each of
which may contain one or more iteration of training and testing. Feature
extraction

Model training can be further broken down into feature extraction, model
selection and feature selection and testing can be performed for model
optimization and

The application of predictive modeling to functional neuroimaging data has been
an active area of research since the early 90s[] and more recently for comparing
connectomes[]. Nearly every analytical method can be appplied in a predictive
modelling framework, but multivariate classification and regression have been
the most commonly used (referred to as MVPA; multivariate prediction analysis).
Multivariate methods are attractive in connectomics because they do not require
any assumptions regarding independence between features and avoid the need to
correct for multiple comparisions since the significance of an entire pattern is
evaluated using a single statistical test. Support vector machines
(classification and regression) are popular for MVPA applications owing to their
proven track record for acheiving good prediction accuracy in high dimensional
datasets (many features) with few observations.

Cross validation

MVPA algorithms typically require the specification of several parameters, the
values of which will impact the quality of the results. These parameters may be
prespecified based on domain specific knowledge, or requirements, determined
using an analytical approach, or optimized using cross validation. In latter
case a nest cross-validation scheme is required to avoid biasing resultant
measures of prediction accuracy. it is necessary to perform a nested
cross-validation These parameters may be set by hand in the case where

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as SVM and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model.  Another approach is to treat feature
selection as a model selection problem in which case the features are
iteratively pruned and a subset is chosen based on prediction performance. A
third approach is to use sparse modeling methods such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed with cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].

>>>>

The quality of a predictive model is determined by its ability to make accurate predictions
from never-before-seen data. This generalization ability is estimated by seperating the 
available data into independet training and testing subsets. A predictive model is learned
from the training data and applied to the testing data to predict corresponding values of the
target variable. Prediction accuracy is then assessed by comparing these predictions to the 
observed value of the target variable. This procedure can be iterated several times for different
splits of the data and averaged to form a single measure of prediction accuracy, in a procedure termed 
cross validation. One of the most important parameters for cross validation is the manner in which data 
are subdivided. There are several strategies for constructing the training dataset ranging from using
all but one observations (leave-one-out cross-validation), using a fraction of the observations
(k-fold cross validation; kfcv), to resampling the observations with replacement (boostrap 
cross-validation, bscv). Although loocv has been popular in neuroimaging applications of MVPA since
it results in the largest amount of training data, its estimates are more variable than 
kfcv or bscv. Split-half cross validation (two-fold cross validation, 2fcv), minimizes the estimation
variance, at the cost of using less training data, and also permits the simultaneous estimation
of model reproducibility using the NPAIRs framework []. Bootstrap cross-validation minimizes estimation
variance while maximising the amount of data available for training at the cost of introducing bias. This
bias is due to the fact that on average 63.2\percent of the testing data overlaps with the training data,
the .632\plus bootstrap corrects for this bias.
No matter which cross validation scheme is used, the accurate estimation of generalization ability
requires that the testing data and training data are independent and identically distributied (iid). The
meaning of indenpdence here depends on the context in which the predictive model will be applied. For 
example, if observations correspond to subjects, then testing with scans acquired from the same subject, 
even acquired at different points in time, will introduce bias. If observations are different time points 
for the same subject, with the goal of testing whether the model can predict a later time point for that 
subject, then scans acquired from a later time point for that subject may be used for testing as long as
enough time has passed to account for temporal autocorrelation.

Another important parameter for estimating generalization ability is the error term employed. For
classification problems error can be simply measured as a '1' if the prediction is correct and a '0' if it
is incorrect, alternatively misclassifications for some groups may be weighted higher then others.
Log-loss or the negative log-likelihood can be used if class probabilities are outputed rather than
simply class membership. For
regression problems, common loss functions are the correlation or mean squared error between predicted and
true target values. The former is preferred in cases when differences in scale and shift are less 
important then the relative ordering of the output balues.  Mean-squared error is preferred when the 
target variable has units and penalizing a model for shift or scale differences is important.

After generalization ability (or error) has been estimated it is compared to what would be acheived by 
``chance'' in order to estimate its significance. In the case of classification problems, the likelihood
of obtaining a generalization accuracy by change can be estimated from a multinomial distribution []. 
Alternatively, a permutation test can be performed in which the target labels are randomly permuted 
several times and the cross-validation error is estimated using each of the permutations. Significance
is the assessed from the fraction of permutations that obtained higher prediction accuracy then was 
obtained from the unpermuted model. This method has the advantage of being available for both 
classification and regression models with the caveat that, any correlationl 
structure between observations to avoid biased results. Null distributions can be constructed for
non-independent data using block permutation or wavelet resampling. When that statistical significance
of the entire model is measured in this manner, the entire pattern is test with a single statistical
test and there is no need to correct for multiple comparisons (unless multiple patterns are tested in 
this manner).

MVPA algorithms typically require the specification of several parameters, the
values of which will impact the quality of the results. These parameters may be
prespecified based on domain specific knowledge, or requirements, determined
using an analytical approach, or optimized using cross validation. In the latter
case a nest cross-validation scheme is required to avoid biasing resultant
measures of prediction accuracy.  Determining the subset of features most relevant 
to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as support vector machines and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model. Another approach is to treat feature
selection as a model selection problem in which case the optimal festure subset 
is chosen based on cross validation based measures of prediction performance [] or 
a measure of their stability[].  A third approach is to use sparse modeling methods 
such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed with cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].
